Bayes Exploration with Brms

### 2023-08-14 Analysis/e1_brms_test.qmd
google sync

shinystan::launch_shinystan(e1_testDistRF2_02)

@@

vx_mm5<- brm(vx ~ 0 + (0+vb|id), data=test,chains=2,silent=0,iter=1200)
mcmc_areas(vx_mm5,prob=.5,regex_pars=c("^r_id\\[1,.*\\]"),regex=T)

mcmc_areas(vx_mm5,prob=.5,regex_pars=c("^r_id\\[2,.*\\]"),regex=T)

mcmc_areas(vx_mm5,prob=.5,regex_pars=c("^r_id\\[3,.*\\]"),regex=T) +
  geom_point(size=.1,alpha=.4)+ggtitle("test")

>>>>

#shinystan::launch_shinystan(e1_testDistRF2_02)

@@

vx_mm5<- brm(vx ~ 0 + (0+vb|id), data=test,chains=2,silent=0,iter=1200)

### 2023-08-14 Misc/Distributional_Explore.qmd
google sync

test |> ggplot(aes(x=dist))+geom_histogram() + facet_wrap(~vb) + ggtitle("empirical_dist")  +
test |> ggplot(aes(x=distS))+geom_histogram() + facet_wrap(~vb) + ggtitle("centered dist")  +
test |> ggplot(aes(x=distS2))+geom_histogram() + facet_wrap(~vb) + ggtitle("scaled dist")

test |> filter(id %in% 1:5) |> ggplot(aes(x=dist))+geom_histogram() + facet_wrap(id~vb) + ggtitle("empirical_dist")

>>>>

# test |> ggplot(aes(x=dist))+geom_histogram() + facet_wrap(~vb) + ggtitle("empirical_dist")  +
# test |> ggplot(aes(x=distS))+geom_histogram() + facet_wrap(~vb) + ggtitle("centered dist")  +
# test |> ggplot(aes(x=distS2))+geom_histogram() + facet_wrap(~vb) + ggtitle("scaled dist")

test |> filter(id %in% 1:5) |> ggplot(aes(x=dist))+geom_histogram() + facet_wrap(id~vb) + ggtitle("empirical_dist")  
plot(density(test$vx))
plot(density(test$dist))

test |> ggplot(aes(x=dist))+geom_density() + facet_wrap(~vb) + ggtitle("empirical_dist") 
test |> ggplot(aes(x=vx))+geom_density() + facet_wrap(~vb) + ggtitle("empirical_dist")

### 2023-08-14 Misc/benchmarks.qmd
google sync

library(ggplot2)
library(brms)

#https://github.com/paul-buerkner/brms/blob/master/vignettes/brms_threading.Rmd

set.seed(54647)
# number of observations
N <- 1E4
# number of group levels
G <- round(N / 10)
# number of predictors
P <- 3
beta <- rnorm(P)
fake <- matrix(rnorm(N * P), ncol = P)
dimnames(fake) <- list(NULL, paste0("x", 1:P))
fake <- transform(
  as.data.frame(fake),
  theta = fake %*% beta,
  g = sample.int(G, N, replace=TRUE)
)
fake  <- merge(fake, data.frame(g = 1:G, eta = rnorm(G)), by = "g")
fake  <- transform(fake, mu = theta + eta)
fake  <- transform(fake, y = rpois(N, exp(mu)))
fake <- fake[sample.int(N, N),]
rownames(fake) <- NULL

kable(head(fake, 10), digits = 3)

# Benchmarks given model with cross-product of tuning parameters CPU
# cores, grainsize and iterations. Models are run with either static
# or non-static scheduler and inits is set by default to 0 on the
# unconstrained scale. Function returns a data-frame with the
# cross-product of the tuning parameters and as result column the
# respective runtime.
benchmark_threading <- function(model, cores = 1, grainsize = 1, iter = 100, 
                                static = FALSE) {

    winfo <- extract_warmup_info(model)
    sims  <- rstan::extract(model$fit)
    init <- list(extract_draw(sims, 1))

    scaling_model <- update(
        model, refresh = 0, 
        threads = threading(1, grainsize = grainsize[1], static = static), 
        chains = 1, iter = 2, backend = "cmdstanr"
    )

    run_benchmark <- function(cores, size, iter) {
        bench_fit <- update(
            scaling_model, warmup=0, iter = iter,
            chains = 1, seed = 1234, inits = init, refresh = 0, save_warmup=TRUE,
            threads = threading(cores, grainsize = size, static = static),
            inv_metric=winfo$inv_metric[[1]],
            step_size=winfo$step_size[[1]],
            adapt_engaged=FALSE
        )
        lf <- sum(subset(nuts_params(bench_fit, inc_warmup=TRUE), Parameter=="n_leapfrog__")$Value)
        elapsed <- sum(colSums(rstan::get_elapsed_time(bench_fit$fit)))

        c(num_leapfrog=lf, runtime=elapsed)
    }

    cases <- expand.grid(cores = cores, grainsize = grainsize, iter = iter)
    res <- with(cases, mapply(run_benchmark, cores, grainsize, iter))
    cbind(cases, as.data.frame(t(res)))
}

benchmark_reference <- function(model, iter=100, inits=0) {
    winfo <- extract_warmup_info(model)
    sims  <- rstan::extract(model$fit)
    init <- list(extract_draw(sims, 1))

    ref_model <- update(
        model, refresh = 0, 
        threads = NULL,
        chains = 1, iter = 2, backend = "cmdstanr"
    )

    run_benchmark_ref <- function(iter_bench) {
        bench_fit <- update(
            ref_model, warmup=0, iter = iter_bench,
            chains = 1, seed = 1234, inits = init, refresh = 0,
            inv_metric=winfo$inv_metric[[1]],
            step_size=winfo$step_size[[1]],
            adapt_engaged=FALSE
        )

        lf <- sum(subset(nuts_params(bench_fit, inc_warmup=TRUE), Parameter=="n_leapfrog__")$Value)
        elapsed <- sum(colSums(rstan::get_elapsed_time(bench_fit$fit)))

        c(num_leapfrog=lf, runtime=elapsed)
    }

    ref <- sapply(iter, run_benchmark_ref)
    ref <- cbind(as.data.frame(t(ref)), iter=iter)
    ref
}

extract_warmup_info <- function(bfit) {
    adapt  <- lapply(rstan::get_adaptation_info(bfit$fit), strsplit, split="\\n")
    step_size  <- lapply(adapt, function(a) as.numeric(strsplit(a[[1]][[1]], " = ")[[1]][2]))
    inv_metric <- lapply(adapt, function(a) as.numeric(strsplit(sub("^# ", "", a[[1]][[3]]), ", ")[[1]]))
    list(step_size=step_size, inv_metric=inv_metric)
}

extract_draw <- function(sims, draw) lapply(sims, abind::asub, idx=draw, dim=1, drop=FALSE)

model_poisson <- brm(
  y ~ 1 + x1 + x2 + (1 | g),
  data = fake,
  family = poisson(),
  iter = 500, # short sampling to speedup example
  chains = 2,
  prior = prior(normal(0,1), class = b) +
    prior(constant(1), class = sd, group = g),
  backend = "cmdstanr",
  threads = threading(4)
)

chunking_bench <- transform(
    data.frame(chunks = 4^(0:3)),
    grainsize = ceiling(N / chunks)
)

iter_test <- c(10, 20, 40)  # very short test runs
scaling_chunking <- benchmark_threading(
  model_poisson,
  cores = 1,                         
  grainsize = chunking_bench$grainsize,  # test various grainsizes
  iter = iter_test,  
  static = TRUE  # with static partitioner
)

# run as reference the model *without* reduce_sum
ref <- benchmark_reference(model_poisson, iter_test)

@@

tf4 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit +(1|bandInt) + (1|id),
  data = test |> filter(id %in% 1:15),
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  threads = threading(n_threads),
  seed=bayes_seed
)
)
tf4

@@

tf2 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit +(1|bandInt) + (1|id),
  data = test |> filter(id %in% 1:15),
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  seed=bayes_seed
)
)
tf2

@@

>>>>

@@

tf4 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit +(1|bandInt) + (1 + bandInt|id),
  data = test |> filter(id %in% 1:15),
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  threads = threading(n_threads),
  seed=bayes_seed
) )
tf4

@@

tf2 <- system.time(
fit_int_norm <- brm(vx ~ 1 + condit +(1|bandInt) + (1 + bandInt|id),
  data = test |> filter(id %in% 1:15),
  family = gaussian(),
  iter=n_iter,
  prior = gprior,
  silent=2,
  cores=n_cores,
  seed=bayes_seed
)
)
tf2

@@

### (20K iterations) - adding condit factor -  2 threads - 4 cores
int only gaussian = 7.78 
 irt version = 45.66 
 int only threading = 8.57 
 irt  threading = 45.97

### (20K iterations) - adding condit factor -  3 threads - 4 cores

int only gaussian = 8.58 
irt version = 47.07 
int only threading = 10.71 
irt  threading = 57.87

### (20K iterations) - condit factor and bandInt RF slope -  2 threads - 4 cores

int only gaussian = 7.85 
 irt version = 80.71 
 int only threading = 8.69 
 irt  threading = 88.3

### 2023-08-14 Misc/scratch/bayes_explore.qmd
google sync

```{r setup}
pacman::p_load(tidyverse,tidybayes,brms,broom,broom.mixed,lme4,emmeans,here,knitr,kableExtra,gt,gghalves,patchwork,ggforce,ggdist)
e1 <- readRDS(here("data/e1_08-04-23.rds"))
source(here("Functions/Display_Functions.R"))
source(here("Functions/org_functions.R"))

@@

 test <- test |> mutate(id=factor(id,levels=unique(id)),
         vxbn=vxb/max(test$vxb),vxn=vx/max(test$vx),bandIntn=bandInt/max(test$vx),bandIntnb=bandInt/max(test$vxb),svx=scale(vx, scale = TRUE),
         vxC=ifelse(vx>2000,2000,vx))

@@

@@

fit_int_norm <- brm(vx ~ 1+bandInt+(1|id), # 2.1 sec execution time with 4 cores
  data = test,
  family = gaussian(),
  prior = c(
    prior(normal(800, 10), class = Intercept),
    prior(normal(400, 10), class = sigma)
  )
)
bayesplot::ppc_dens_overlay_grouped(test$vx,
  yrep = posterior_predict(fit_int_norm, ndraws = 100),
  group = test$bandInt
)

>>>>

```{r setup}
pacman::p_load(tidyverse,tidybayes,bayesplot,brms,broom,broom.mixed,lme4,emmeans,here,knitr,kableExtra,gt,gghalves,patchwork,ggforce,ggdist)
e1 <- readRDS(here("data/e1_08-04-23.rds"))
source(here("Functions/Display_Functions.R"))
source(here("Functions/org_functions.R"))

@@

 test <- test |> mutate(id=factor(id,levels=unique(id)),
         vxbn=vxb/max(test$vxb),vxn=vx/max(test$vx),
         bandIntn=bandInt/max(test$vx),bandIntnb=bandInt/max(test$vxb),
         svx=scale(vx, scale = TRUE),
         vxC=ifelse(vx>2000,2000,vx))

@@

options(mc.cores = 4, brms.backend = "cmdstanr")
```

@@

fit_int_norm <- brm(vx ~ 1+bandInt+(1|id), # 2.1 sec execution time with 4 cores
  data = test,
  family = gaussian(),
  prior = c(
    prior(normal(800, 10), class = Intercept),
    prior(normal(400, 10), class = sigma)
  )
)

bayesplot::ppc_dens_overlay_grouped(test$vx,
  yrep = posterior_predict(fit_int_norm, ndraws = 100),
  group = test$bandInt
)
